### 2020-06-29 TCS_version/01-introduction.tex
qualche progresso

%Formally, Reaction Systems are defined as a rewriting formalism.
%In this paper, we present an encoding from RSs, 
%to the  open multiparty process algebra c\CNA,\footnote{After `chained Core Network 
%Algebra'.} a variant of the {\tt link}-calculus~\cite{BodeiBB12,BBB17} without name mobility.
In this paper, which is an extended version of~\cite{BBF19},
we present an encoding
%encoding 
from RSs
to the chained Core Network  Algebra (c\CNA), a variant of the 
 open multiparty process algebra~\CNA~\cite{BBB17}; the \CNA~equipped with mobility is referred to as the {\tt link}-calculus~\cite{BodeiBB12,BODEI2020104587}.
 Here mobility is not needed.
This formalism allows several processes to synchronise and 
communicate altogether, at the same time, with a new interaction
mechanism based on links and link chains. 
The initial motivation for introducing
an open multiparty mechanism in~\cite{BodeiBB12}
% this mechanism 
was to encode
Mobile Ambients~\cite{CardelliG00}, getting a much stronger operational
correspondence than any available in the literature, such as the one in~\cite{B16}.
Later it was shown that the {\tt link}-calculus allowed one to easily encode calculi for biology equipped with membranes, as in~\cite{BodeiBBC14}.

>>>>

%Formally, Reaction Systems are defined as a rewriting formalism.
%In this paper, we present an encoding from RSs, 
%to the  open multiparty process algebra c\CNA,\footnote{After `chained Core Network 
%Algebra'.} a variant of the {\tt link}-calculus~\cite{BodeiBB12,BBB17} without name mobility.
In this paper, which is an extended version of~\cite{BBF19},
we present an encoding
%encoding 
from RSs
to the chained Core Network  Algebra (c\CNA), a variant of the 
 open multiparty process algebra~\CNA~\cite{BBB17}; the \CNA~equipped with mobility is referred to as the {\tt link}-calculus~\cite{BodeiBB12,BODEI2020104587}.
 Here mobility is not needed.
This formalism allows several processes to synchronise and 
communicate altogether, at the same time, with a new interaction
mechanism based on links and link chains. 
The initial motivation for introducing
an open multiparty mechanism in~\cite{BodeiBB12}
% this mechanism 
was to encode
Mobile Ambients~\cite{CardelliG00}, getting a much stronger operational
correspondence than any available in the literature, such as the ones in~\cite{CZ07,B16}.
Later it was shown that the {\tt link}-calculus allowed one to easily encode calculi for biology equipped with membranes, as in~\cite{BodeiBBC14}.

### 2020-06-29 TCS_version/02-ReactionSystem.tex
qualche progresso

The sets 
$R, I, P$ are also written $R_a, I_a, P_a$ and called the 
\emph{reactant set} of $a$, the  
\emph{inhibitor set} of $a$, and the \emph{product set} of $a$, respectively. 
All reactants are needed for the reaction to take place.
Any inhibitor blocks the reaction if it is present. Products are the outcome of the reaction.
Also,  $R_a \cup I_a$ is the set of the resources of $a$ and $ rac(S)$
denotes the set of all reactions in $S$.
%
Because $R$ and $I$ are non empty, all products are produced from at least one reactant and every reaction can be inhibited in some way. 
Sometimes artificial inhibitors are used that are never produced by any reaction.
For the sake of simplicity, in some examples, we will allow $I$ to be empty.
%Note that since $R_a$ and $I_a$ are both nonempty, we always 
%have $|M_a| \geq 2$.

\begin{definition}[Reaction System]
A Reaction System (RS) is an ordered pair ${\cal A} = (S, A)$ 
such that $S$ is a finite set of  entities, and $A \subseteq rac(S)$ is a set of reactions over $S$.
%The set $S$ is called the background (set) of $A$.\\
%The main notion concerning the functioning of a reaction system is defined 
%as follows.
\end{definition}

>>>>

The sets 
$R, I, P$ are also written $R_a, I_a, P_a$ and called the 
\emph{reactant set} of $a$, the  
\emph{inhibitor set} of $a$, and the \emph{product set} of $a$, respectively. 
All reactants are needed for the reaction to take place.
Any inhibitor blocks the reaction if it is present. Products are the outcome of the reaction.
Also,  $R_a \cup I_a$ is the set of the resources of $a$ and $\mathit{rac}(S)$
denotes the set of all reactions in $S$.
%
Because $R$ and $I$ are non empty, all products are produced from at least one reactant and every reaction can be inhibited in some way. 
Sometimes artificial inhibitors are used that are never produced by any reaction.
For the sake of simplicity, in some examples, we will allow $I$ to be empty.
%Note that since $R_a$ and $I_a$ are both nonempty, we always 
%have $|M_a| \geq 2$.

\begin{definition}[Reaction System]
A Reaction System (RS) is an ordered pair ${\cal A} = (S, A)$ 
such that $S$ is a finite set of  entities, and $A \subseteq \mathit{rac}(S)$ is a set of reactions over $S$.
%The set $S$ is called the background (set) of $A$.\\
%The main notion concerning the functioning of a reaction system is defined 
%as follows.
\end{definition}

### 2020-06-29 TCS_version/04-CCNA.tex
qualche progresso

In this section we introduce  the syntax and operational semantics
of the process algebra c\CNA\  (chained \CNA) to be used for encoding RSs.
As already explained in the Introduction, c\CNA\ is a variant of \CNA~\cite{BBB17}, the non-mobile
fragment of {\tt link}-calculus~\cite{BodeiBB12,BODEI2020104587}.
In c\CNA\  the action prefixes are link chains and not just links.
%In this section we just focus on the basic ideas of link
% interaction, while we shall enhance the model with name mobility in
% the next section.

\paragraph{Link Chains}
Let $\mathcal{C}$ be the set of channels, ranged over by $a,b,...$, and 
let $\mathcal{A} = \mathcal{C} \cup \setof{\silent} \cup \setof{{\noacts}}$ be the set of actions, ranged over by $\alpha,\beta,...$,
where the symbol $\silent$ denotes a \emph{silent} action, while the symbol $\noact$ denotes a \emph{virtual} (non-specified) action.
A \emph{link} is a pair $\ell=\link{\alpha}{\beta}$;
it is \emph{solid} if $\alpha,\beta \neq \ \noacts$; 
{\color{red} $\alpha$ and $\beta$ are two interaction points, to fix the idea $\alpha$ is the action, and $\beta$ is the coaction.}
The link $\link{\noact}{\noact}$ is called \emph{virtual}.
A link is \emph{valid} if it is solid or virtual.
We let $\mathcal{L}$ be the set of valid links.
%
A \emph{link chain} is a 
finite sequence $v = \ell_{1}...\ell_{n}$ of (valid) links  $\ell_{i} = \link{\alpha_{i}}{\beta_{i}}$ such that:
\begin{enumerate}
\item for any $i\in [1,n-1]$, 
$\left\{\begin{array}{ll}
\beta_{i},\alpha_{i+1}\in \mathcal{C} & \mbox{ implies } \beta_{i} = \alpha_{i+1}\\
\beta_{i}=\silent & \mbox{ iff } \alpha_{i+1}=\silent
\end{array}\right.
$
\item  $\exists i \in [1,n]. \ \ell_i \neq \link{\noact}{\noact}$.
\end{enumerate}

Virtual links represent missing elements of a chain. 
A chain is called \emph{solid} if it does not contain any virtual link.
The equivalence $\blackstretcheq$ models expansion/contraction of virtual links to adjust the length of a link chain.

\begin{definition}[Equivalence $\blackstretcheq$]\label{def:black}
We let $\blackstretcheq$ be the least equivalence relation 
over link chains closed under the axioms (whenever both sides are well defined):
\[
\begin{array}{rclcrcl}
v\link{\noact}{\noact} & \blackstretcheq &  v & \qquad &
v_1 \startchain{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}v_2 & \blackstretcheq & v_1 \link{\noact}{\noact} v_2\\
\link{\noact}{\noact}v & \blackstretcheq & v & &
v_1 \startchain{\alpha}\chainedlink{a}{a}\chainend{\beta} v_2
& \blackstretcheq & 
v_1 \startchain{\alpha}\chainedlink{a}{\noact}\chainedlink{\noact}{a}\chainend{\beta}v_2 
\end{array}
\]
\end{definition}

Two link chains of equal length can be merged whenever each position occupied by a solid link in one chain is occupied by a virtual link in the other chain and solid links in adjacent positions match. Positions occupied by virtual links in both chains remain virtual. Merging is denoted by $\merges{v_1}{v_2}$.
%
For example, given $v_1 = \startchain{\silent}\chainedlink{a}{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}$, $v_2 = \startchain{\noact}\chainedlink{\noact}{a}\chainedlink{b}{\noact}\chainend{\noact}$ and $v = \startchain{\silent}\chainedlink{a}{a}\chainedlink{b}{\noact}\chainend{\noact}$
we have $\merges{v_1}{v_2} = v$, whereas $\merges{v_1}{v}$ is not defined. Notably the merge operation is commutative and associative.

Some names in a link chain can be restricted as non observable and transformed into silent actions $\silent$. This is possible only if they are matched by some adjacent link. Restriction is denoted by $\restrict{a}v$.
%
For example, given $v = \startchain{\silent}\chainedlink{a}{a}\chainedlink{b}{\noact}\chainend{\noact}$ as above, we have 
$\restrict{a}v = \startchain{\silent}\chainedlink{\silent}{\silent}\chainedlink{b}{\noact}\chainend{\noact}$,
whereas $\restrict{b}v$ is not defined.

\paragraph{Syntax}
%\begin{definition}
The set of c\CNA\ processes, denoted as $\cal{P}$ and ranged over by $P,Q$, is defined by the following grammar:
\[
\begin{array}{rcl}
P,Q & ::= &
%\nil \prodsep
%X \prodsep
%s.P  \prodsep
\sum_{i\in I} \upsilon_i.P_i \prodsep
P|Q \prodsep
\restrict{a} P \prodsep
P[\phi] \prodsep
A
\end{array}
\]
\noindent
where $\upsilon_i$ is a  link chain,
%(i.e.\  $\upsilon_i =
%\startchain{\alpha}\chainedlink{\beta}{\noacts}\chainedlink{\noacts}{\alpha'}\chainend{\beta'}$
%with $\alpha,\alpha',\beta,\beta'\neq \ \noacts$),
$\phi$ is a channel renaming function,
and $A$ is a process identifier for which we assume a definition $A
\defeq P$ is available in a given set $\Delta$ of (possibly
recursive) process definitions. We let $\nil$, the inactive process,
denote the empty summation.
%\end{definition}

The syntax of c\CNA\  extends that of \CNA~\cite{BBB17} by allowing to use link chains as prefixes instead of links.
For the rest it features nondeterministic (guarded) choice, parallel composition, restriction, relabelling and possibly recursive definitions. Here we do not consider name mobility, which is present instead in the {\tt link}-calculus.

@@

The operational semantics of c\CNA\ is defined in the SOS style by the inference rules in Fig.\ref{fig:cnasos}. The rules are reminiscent of those for Milner's CCS and they essentially coincide with those of \CNA\ in~\cite{BBB17}. The only difference is due to the presence of prefixes that are link chains.
Briefly: rule (\textit{Sum}) selects one alternative and puts as label a possible contraction/expansion of the link chain in the selected prefix; rule (\textit{Ide}) selects one transition of the defining process for a constant; rule (\textit{Res}) restricts some names in the label (it cannot be applied when $\restrict{a}v$ is not defined); rules (\textit{Lpar}) and (\textit{Rpar}) account for interleaving in parallel composition; rule (\textit{Com}) 
synchronises interactions (it cannot be applied when $\merges{v}{v'}$ is not defined).

\begin{example}
As some simple examples, consider the recursive process definitions $H \defeq \link{a}{b}.\link{a}{c}.H$ and $K \defeq \link{a}{b}.K + \link{a}{c}.K$: the former is some sort of forwarder that alternates between providing a link from $a$ to $b$ and to $c$; the latter is a nondeterministic forwarder that at each step can link $a$ to $b$ or to $c$. 
\end{example}

@@

\begin{figure}[t]
\begin{center} 
\begin{prooftree} 
\AxiomC{$\upsilon \blackstretcheq \upsilon_j\quad j\in I$} %\mbox{($\ell$ {\em only} solid link in $s$)}
\RightLabel{\scriptsize(\textit{Sum})}
\UnaryInfC{$\sum_{i\in I} \upsilon_i.P_i \xrightarrow{\upsilon} P_j$} 
\DisplayProof
\
\AxiomC{$P \xrightarrow{\upsilon} P'$}
 \AxiomC{$(A \defeq P)\in\Delta$}
\RightLabel{\scriptsize(\textit{Ide})}
\BinaryInfC{$A \xrightarrow{\upsilon} P'$} 
%\AxiomC{$P \xrightarrow{v} P'$} 
%\RightLabel{\scriptsize(\textit{Lsum})}
%\UnaryInfC{$P+Q \xrightarrow{v} P'$} 
%\DisplayProof
%\
%\AxiomC{$Q \xrightarrow{s} Q'$} 
%\RightLabel{\scriptsize(Rsum)}
%\UnaryInfC{$P+Q \xrightarrow{s} Q'$} 
\end{prooftree} 
\end{center}

\begin{center} 
\begin{prooftree} 
\AxiomC{$P \xrightarrow{\upsilon} P'$} 
\RightLabel{\scriptsize(\textit{Res})}
\UnaryInfC{$\restrict{a}P \xrightarrow{\restrict{a}\upsilon} \restrict{a}P'$} 
\DisplayProof
%\,
%\AxiomC{$P \xrightarrow{v} P'$} 
%\RightLabel{\scriptsize(\textit{Rel})}
%\UnaryInfC{$P[\phi] \xrightarrow{v[\phi]} P'[\phi]$} 
%\DisplayProof
\AxiomC{$P \xrightarrow{v} P'$} 
\RightLabel{\scriptsize(\textit{Lpar})}
\UnaryInfC{$P|Q \xrightarrow{\upsilon} P'|Q$} 
\DisplayProof
%\,
%\AxiomC{$Q \xrightarrow{s} Q'$} 
%\RightLabel{\scriptsize(Rpar)}
%\UnaryInfC{$P|Q \xrightarrow{s} P|Q'$} 
%\end{prooftree} 
%\end{center}
%\begin{center} 
%\begin{prooftree} 
\AxiomC{$P \xrightarrow{\upsilon'} P'$} 
\AxiomC{$Q \xrightarrow{\upsilon} Q'$} 
\RightLabel{\scriptsize(\textit{Com})}
\BinaryInfC{$P|Q \xrightarrow{\merges{\upsilon}{\upsilon'}} P'|Q'$} 
%\DisplayProof
\end{prooftree} 
\end{center}
\caption{SOS semantics of the c\CNA\  (rules (\textit{Rel}) and (\textit{Rpar}) omitted).}
\label{fig:cnasos}
\end{figure}

@@

Hereafter we make use of some new notations for link chains that will facilitate  
the presentation of our 
%translation.
{\color{red} embedding.}

\begin{definition}[Replication]
Let $\upsilon$ be a link chain such that $\upsilon\upsilon$ is also a valid link chain. Its $n$ times replication $\upsilon^n$ is defined recursively by letting
$\upsilon^0 = \epsilon$ (i.e. the empty chain) and $\upsilon^n = \upsilon\upsilon^{n-1}$.
%\[
%\begin{array}{c@{\qquad}c}
%%\ell ^1 = \ell & \ell^n = \ell^{n-1}\ell\\
%\upsilon^1 = \upsilon& \upsilon^n = \upsilon^{n-1}\upsilon
%\end{array}
%\]
%with the hypothesis that all the links in the resulting link chains match.
%, and that $\upsilon^0 = \epsilon$ is the empty chain.
\end{definition}

For example,  the expression $(\startchain{a}\chainedlink{b}{\noact}\chainend{\noact})^3$  denotes the chain $\startchain{a}\chainedlink{b}{\noact}\chainedlink{\noact}{a}\chainedlink{b}{\noact}\chainedlink{\noact}{a}\chainedlink{b}{\noact}\chainend{\noact}$. 
Instead the expression $(\link{a}{b})^2$ is ill-defined because $a$ does not match with $b$.

@@

\begin{definition}[Half links]
Let $a$ be a channel name, we define the \emph{half left link}  $\link{a}{}$,   and the  \emph{half right link} $\link{}{a}$.
\end{definition}

\begin{definition}[Open block]
Let $R$ be a set of names. We 
define an \emph{open block} as $\blockchain{a_i}{a_o}{a \in R}$, where $a_i$ and $a_o$ are annotated version of the name $a$, as
\[
\begin{array}{l@{\quad}l@{\quad}l}
\mbox{set}& \mbox{block of chain}& \mbox{result}\\
R= \emptyset & \blockchain{a_i}{a_o}{a \in R}   &\epsilon\\
R = \{ b\}    &\blockchain{a_i}{a_o}{a \in R}  & {}_{b_i}^{\noact}\chainedlink{\noact}{b_o}\\
R = \{b\} \cup R' & \blockchain{a_i}{a_o}{a \in R} & {}_{b_i}^{\noact}\chainedlink{\noact}{b_o}\link{}{}\blockchain{a_i}{a_o}{a \in R'}
\end{array}
\]
\end{definition}
%Please note that, channel names of juxtaposed open blocks of chain are not required to match.

%\noindent
We then combine half links and  open blocks to form regular link chains.
%
%  $\halfl{a}\blockchain{c_i}{c_o}{c \in R}\halfr{b}$ to form a regular link chain; or we can put together half links and open blocks of chain:  $\halfl{a}\blockchain{c_i}{c_o}{c \in R}\blockchain{c_i}{c_o}{c \in R'}\dots \blockchain{c_i}{c_o}{c \in R^{''}}\halfr{b}$ to still form a regular  link chain.
% Again,  all the channel names appearing in the half links or in the open blocks of chain are not required to match.
 For example, for $R = \{a,b\}$ the expression $ \blockchain{c_i}{c_o}{c \in R} $
denotes the block ${}_{a_i}^{\noact}\chainedlink{\noact}{a_o} \backslash{}_{b_i}^{\noact}\chainedlink{\noact}{b_o}$; and the expression $\link{r_1}{} \blockchain{c_i}{c_o}{c \in R}\link{}{r_2} $ denotes the chain 
$ \startchain{r_1}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}\chainedlink{b_i}{\noact}\chainedlink{\noact}{b_o}\chainend{r_2}$.

>>>>

In this section we introduce  the syntax and operational semantics
of the process algebra c\CNA\  (chained \CNA)~\cite{BBF19} to be used for encoding RSs.
As already explained in the Introduction, c\CNA\ is a variant of \CNA~\cite{BBB17}, the non-mobile
fragment of {\tt link}-calculus~\cite{BodeiBB12,BODEI2020104587}.
In c\CNA\  the action prefixes are link chains and not just links.
%In this section we just focus on the basic ideas of link
% interaction, while we shall enhance the model with name mobility in
% the next section.

\paragraph{Link Chains}
Let $\mathcal{C}$ be the set of channels, ranged over by $\mathsf{a},\mathsf{b},...$, and 
let $\mathit{Act} \defeq \mathcal{C} \cup \setof{\silent} \cup \setof{{\noacts}}$ be the set of actions, ranged over by $\alpha,\beta,...$,
where the symbol $\silent$ denotes a \emph{silent} action, while the symbol $\noact$ denotes a \emph{virtual} (non-specified) action.
A \emph{link} is a pair $\ell=\link{\alpha}{\beta}$;
it is \emph{solid} if $\alpha,\beta \neq \ \noacts$; 
intuitively, $\alpha$ and $\beta$ are two interaction points, one for incoming requests and the other for outgoing requests.
The link $\link{\noact}{\noact}$ is called \emph{virtual}.
A link is \emph{valid} if it is solid or virtual.
We let $\mathcal{L}$ be the set of valid links.
%
A \emph{link chain} is a 
finite sequence $v = \ell_{1}...\ell_{n}$ of (valid) links  $\ell_{i} = \link{\alpha_{i}}{\beta_{i}}$ such that:
\begin{enumerate}
\item for any $i\in [1,n-1]$, 
$\left\{\begin{array}{ll}
\beta_{i},\alpha_{i+1}\in \mathcal{C} & \mbox{ implies } \beta_{i} = \alpha_{i+1}\\
\beta_{i}=\silent & \mbox{ iff } \alpha_{i+1}=\silent
\end{array}\right.
$
\item  $\exists i \in [1,n]. \ \ell_i \neq \link{\noact}{\noact}$.
\end{enumerate}

Virtual links represent missing elements of a chain. 
A chain is called \emph{solid} if it does not contain any virtual link.
The empty chain is denoted by $\epsilon$.
The equivalence $\blackstretcheq$ models expansion/contraction of virtual links to adjust the length of a link chain.

\begin{definition}[Equivalence $\blackstretcheq$]\label{def:black}
We let $\blackstretcheq$ be the least equivalence relation 
over link chains closed under the axioms (whenever both sides are well defined):
\[
\begin{array}{rclcrcl}
v\link{\noact}{\noact} & \blackstretcheq &  v & \qquad &
v_1 \startchain{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}v_2 & \blackstretcheq & v_1 \link{\noact}{\noact} v_2\\[5pt]
\link{\noact}{\noact}v & \blackstretcheq & v & &
v_1 \startchain{\alpha}\chainedlink{\mathsf{a}}{\mathsf{a}}\chainend{\beta} v_2
& \blackstretcheq & 
v_1 \startchain{\alpha}\chainedlink{\mathsf{a}}{\noact}\chainedlink{\noact}{\mathsf{a}}\chainend{\beta}v_2 
\end{array}
\]
\end{definition}

Two link chains of equal length can be merged whenever each position occupied by a solid link in one chain is occupied by a virtual link in the other chain and solid links in adjacent positions match. Positions occupied by virtual links in both chains remain virtual. Merging is denoted by $\merges{v_1}{v_2}$.
%
For example, given $v_1 = \startchain{\silent}\chainedlink{\mathsf{a}}{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}$, $v_2 = \startchain{\noact}\chainedlink{\noact}{\mathsf{a}}\chainedlink{\mathsf{b}}{\noact}\chainend{\noact}$ and $v = \startchain{\silent}\chainedlink{\mathsf{a}}{\mathsf{a}}\chainedlink{\mathsf{b}}{\noact}\chainend{\noact}$
we have $\merges{v_1}{v_2} = v$, whereas $\merges{v_1}{v}$ is not defined. Notably the merge operation is commutative and associative.

Some names in a link chain can be restricted as non observable and transformed into silent actions $\silent$. This is possible only if they are matched by some adjacent link. Restriction is denoted by $\restrict{\mathsf{a}}v$.
%
For example, given $v = \startchain{\silent}\chainedlink{\mathsf{a}}{\mathsf{a}}\chainedlink{\mathsf{b}}{\noact}\chainend{\noact}$ as above, we have 
$\restrict{\mathsf{a}}v = \startchain{\silent}\chainedlink{\silent}{\silent}\chainedlink{\mathsf{b}}{\noact}\chainend{\noact}$,
whereas $\restrict{\mathsf{b}}v$ is not defined.

\paragraph{Syntax}
%\begin{definition}
The set of c\CNA\ processes, denoted as $\cal{P}$ and ranged over by $P,Q$, is defined by the following grammar:
\[
\begin{array}{rcl}
P,Q & ::= &
\nil \prodsep
%X \prodsep
\upsilon.P  \prodsep
%\sum_{i\in I} \upsilon_i.P_i \prodsep
P+Q \prodsep
P|Q \prodsep
\restrict{\mathsf{a}} P \prodsep
%P[\phi] \prodsep
A
\end{array}
\]
\noindent
where $\upsilon_i$ is a  link chain,
%(i.e.\  $\upsilon_i =
%\startchain{\alpha}\chainedlink{\beta}{\noacts}\chainedlink{\noacts}{\alpha'}\chainend{\beta'}$
%with $\alpha,\alpha',\beta,\beta'\neq \ \noacts$),
%$\phi$ is a channel renaming function,
and $A$ is a process identifier.
The syntax of c\CNA\  extends that of \CNA~\cite{BBB17} by allowing to use link chains as prefixes instead of links, i.e. we allow to write $\upsilon.P$ instead of $\ell.P$.
For the rest it features nondeterministic choice $P+Q$ (also called sum), parallel composition $P|Q$, restriction $\restrict{\mathsf{a}} P$, and possibly recursively defined process identifiers $A$. 
Here we do not consider name mobility, which is present instead in the {\tt link}-calculus and we omit the relabelling operator of \CNA\ (not needed in the encoding).

As common in process algebras we restrict to consider prefix-guarded sums $\upsilon_1.P_1 + \upsilon_2.P_2$ and, exploiting associativity, we use the shorthand $\sum_{i\in I} \upsilon_i.P_i$ for a finite set of indexes $I=\{i_1,...,i_k\}$ instead of $\upsilon_{i_1}.P_{i_1} + \cdots + \upsilon_{i_k}.P_{i_k}$.
The inactive process $\nil$ is thus just the empty summation.

Regarding process constants,  we rely on a given set $\Delta=\{A_i\defeq P_i\}_{i\in I}$ of (possibly
recursive) process definitions.
%\end{definition}

@@

The operational semantics of c\CNA\ is defined in the SOS style by the inference rules in Fig.\ref{fig:cnasos}. The rules are reminiscent of those for Milner's CCS and they essentially coincide with those of \CNA\ in~\cite{BBB17}. The only difference is due to the presence of prefixes that are link chains.
Briefly: rule (\textit{Sum}) selects one alternative and puts as label a possible contraction/expansion of the link chain in the selected prefix; rule (\textit{Ide}) selects one transition of the defining process for a constant; rule (\textit{Res}) restricts some names in the label (it cannot be applied when $\restrict{\mathsf{a}}v$ is not defined); rules (\textit{Lpar}) and (\textit{Rpar}) account for interleaving in parallel composition; rule (\textit{Com}) 
synchronises interactions (it cannot be applied when $\merges{v}{v'}$ is not defined).

\begin{example}
As some simple examples, consider the recursive process definitions $H \defeq \link{\mathsf{a}}{\mathsf{b}}.\link{\mathsf{a}}{\mathsf{c}}.H$ and $K \defeq \link{\mathsf{a}}{\mathsf{b}}.K + \link{\mathsf{a}}{\mathsf{c}}.K$: the former recursively provides a link from $\mathsf{a}$ to $\mathsf{b}$ and then, at the next step, from $\mathsf{a}$ to $\mathsf{c}$; the latter provides at each step a link from $\mathsf{a}$ to $\mathsf{b}$ or from $\mathsf{a}$ to $\mathsf{c}$, nondeterministically. 
\end{example}

@@

\begin{figure}[t]
\begin{center} 
\begin{prooftree} 
\AxiomC{$\upsilon \blackstretcheq \upsilon_j\quad j\in I$} %\mbox{($\ell$ {\em only} solid link in $s$)}
\RightLabel{\scriptsize(\textit{Sum})}
\UnaryInfC{$\sum_{i\in I} \upsilon_i.P_i \xrightarrow{\upsilon} P_j$} 
\DisplayProof
\qquad
\AxiomC{$P \xrightarrow{\upsilon} P'$}
 \AxiomC{$(A \defeq P)\in\Delta$}
\RightLabel{\scriptsize(\textit{Ide})}
\BinaryInfC{$A \xrightarrow{\upsilon} P'$} 
%\AxiomC{$P \xrightarrow{v} P'$} 
%\RightLabel{\scriptsize(\textit{Lsum})}
%\UnaryInfC{$P+Q \xrightarrow{v} P'$} 
%\DisplayProof
%\
%\AxiomC{$Q \xrightarrow{s} Q'$} 
%\RightLabel{\scriptsize(Rsum)}
%\UnaryInfC{$P+Q \xrightarrow{s} Q'$} 
\end{prooftree} 
\end{center}

\begin{center} 
\begin{prooftree} 
\AxiomC{$P \xrightarrow{\upsilon} P'$} 
\RightLabel{\scriptsize(\textit{Res})}
\UnaryInfC{$\restrict{\mathsf{a}}P \xrightarrow{\restrict{\mathsf{a}}\upsilon} \restrict{a}P'$} 
\DisplayProof
\quad
%\,
%\AxiomC{$P \xrightarrow{v} P'$} 
%\RightLabel{\scriptsize(\textit{Rel})}
%\UnaryInfC{$P[\phi] \xrightarrow{v[\phi]} P'[\phi]$} 
%\DisplayProof
\AxiomC{$P \xrightarrow{v} P'$} 
\RightLabel{\scriptsize(\textit{Lpar})}
\UnaryInfC{$P|Q \xrightarrow{\upsilon} P'|Q$} 
\DisplayProof
\quad
\AxiomC{$Q \xrightarrow{v} Q'$} 
\RightLabel{\scriptsize(\textit{Rpar})}
\UnaryInfC{$P|Q \xrightarrow{\upsilon} P|Q'$} 
%\DisplayProof
\end{prooftree} 
\end{center}

\begin{center} 
\begin{prooftree} 
%\,
%\AxiomC{$Q \xrightarrow{s} Q'$} 
%\RightLabel{\scriptsize(Rpar)}
%\UnaryInfC{$P|Q \xrightarrow{s} P|Q'$} 
%\end{prooftree} 
%\end{center}
%\begin{center} 
%\begin{prooftree} 
\AxiomC{$P \xrightarrow{\upsilon'} P'$} 
\AxiomC{$Q \xrightarrow{\upsilon} Q'$} 
\RightLabel{\scriptsize(\textit{Com})}
\BinaryInfC{$P|Q \xrightarrow{\merges{\upsilon}{\upsilon'}} P'|Q'$} 
%\DisplayProof
\end{prooftree} 
\end{center}
\caption{SOS semantics of c\CNA\  processes.}
\label{fig:cnasos}
\end{figure}

@@

Hereafter we make use of some new notations for link chains that will facilitate  
the presentation of our 
%translation.
encoding.

\begin{definition}[Replication]
Let $\upsilon$ be a valid link chain such that $\upsilon\upsilon$ is also a valid link chain. The $n$ times replication of $\upsilon$, written $\upsilon^n$, is defined recursively by letting
$\upsilon^0 = \epsilon$ (i.e. the empty chain) and $\upsilon^n = \upsilon\upsilon^{n-1}$.
%\[
%\begin{array}{c@{\qquad}c}
%%\ell ^1 = \ell & \ell^n = \ell^{n-1}\ell\\
%\upsilon^1 = \upsilon& \upsilon^n = \upsilon^{n-1}\upsilon
%\end{array}
%\]
%with the hypothesis that all the links in the resulting link chains match.
%, and that $\upsilon^0 = \epsilon$ is the empty chain.
\end{definition}

For example,  the expression $(\startchain{\mathsf{a}}\chainedlink{\mathsf{b}}{\noact}\chainend{\noact})^3$  denotes the chain $\startchain{\mathsf{a}}\chainedlink{\mathsf{b}}{\noact}\chainedlink{\noact}{\mathsf{a}}\chainedlink{\mathsf{b}}{\noact}\chainedlink{\noact}{\mathsf{a}}\chainedlink{\mathsf{b}}{\noact}\chainend{\noact}$. 
Instead the expression $(\link{\mathsf{a}}{\mathsf{b}})^2$ is ill-defined because $\mathsf{a}$ does not match with $\mathsf{b}$.

@@

\begin{definition}[Half links]
Let $\mathsf{a}$ be a channel name, we denote by $\link{\mathsf{a}}{}$ the \emph{half left link},  and by $\link{}{\mathsf{a}}$ the  \emph{half right link}.
\end{definition}

In the encoding of RS we will make extensive use of subscripted names:  each name $\mathsf{a}$ will come in two variants $\mathsf{a}_i$ and $\mathsf{a}_o$. This is just a technical issue to prevent accidental matching between links. To see why this is important, compare the chain 
$\startchain{\silent}\chainedlink{\mathsf{a}}{\noact}\chainedlink{\noact}{\mathsf{a}}\chainend{\silent}$ with 
$\startchain{\silent}\chainedlink{\mathsf{a}_i}{\noact}\chainedlink{\noact}{\mathsf{a}_o}\chainend{\silent}$: the former is $\blackstretcheq$ equivalent to the solid chain $\startchain{\silent}\chainedlink{\mathsf{a}}{\mathsf{a}}\chainend{\silent}$; the latter cannot become solid unless merged with a chain that links $\mathsf{a}_i$ to $\mathsf{a}_o$, like 
$\startchain{\noact}\chainedlink{\noact}{\mathsf{a}_i}\chainedlink{\mathsf{a}_o}{\noact}\chainend{\noact}$.
This form of subscripting is exploited in the definition of open blocks.

\begin{definition}[Open block]
Let $\sigma$ be a finite sequence of names. We 
define an \emph{open block} as $\blockchain{\mathsf{a}_i}{\mathsf{a}_o}{\mathsf{a} \in \sigma}$, where $\mathsf{a}_i$ and $\mathsf{a}_o$ are annotated version of the name $\mathsf{a}$ (as explained above), by letting
\[
\blockchain{\mathsf{a}_i}{\mathsf{a}_o}{\mathsf{a} \in \sigma} \defeq
\left\{
\begin{array}{ll}
\epsilon 
& \mbox{if $\sigma=\epsilon$ is the empty sequence}\\[5pt]
{}_{\mathsf{b}_i}^{\noact}\chainedlink{\noact}{\mathsf{b}_o}\blockchain{\mathsf{a}_i}{\mathsf{a}_o}{\mathsf{a} \in \rho}
& \mbox{if $\sigma=\mathsf{b}\rho$}
\end{array}
\right.
\]
%\[
%\begin{array}{l@{\quad}l@{\quad}l}
%\mbox{set}& \mbox{block of chain}& \mbox{result}\\
%R= \emptyset & \blockchain{a_i}{a_o}{a \in R}   &\epsilon\\
%R = \{ b\}    &\blockchain{a_i}{a_o}{a \in R}  & {}_{b_i}^{\noact}\chainedlink{\noact}{b_o}\\
%R = \{b\} \cup R' & \blockchain{a_i}{a_o}{a \in R} & {}_{b_i}^{\noact}\chainedlink{\noact}{b_o}\link{}{}\blockchain{a_i}{a_o}{a \in R'}
%\end{array}
%\]
\end{definition}
%Please note that, channel names of juxtaposed open blocks of chain are not required to match.

Abusing the notation, we will use the open block notation for sets of names rather than sequences, assuming the names in the set are taken  according to some default order (e.g. the lexicographic one).

%\noindent
We then combine half links and  open blocks to form valid link chains.
%
%  $\halfl{a}\blockchain{c_i}{c_o}{c \in R}\halfr{b}$ to form a regular link chain; or we can put together half links and open blocks of chain:  $\halfl{a}\blockchain{c_i}{c_o}{c \in R}\blockchain{c_i}{c_o}{c \in R'}\dots \blockchain{c_i}{c_o}{c \in R^{''}}\halfr{b}$ to still form a regular  link chain.
% Again,  all the channel names appearing in the half links or in the open blocks of chain are not required to match.

 For example, for $X = \{\mathsf{a},\mathsf{b}\}$ the expression $ \blockchain{\mathsf{c}_i}{\mathsf{c}_o}{\mathsf{c} \in X} $
denotes the block ${}_{\mathsf{a}_i}^{\noact}\chainedlink{\noact}{\mathsf{a}_o} \backslash{}_{\mathsf{b}_i}^{\noact}\chainedlink{\noact}{\mathsf{b}_o}$; and the expression $\link{\mathsf{r}_1}{} \blockchain{\mathsf{c}_i}{\mathsf{c}_o}{\mathsf{c} \in X}\link{}{\mathsf{r}_2} $ denotes the chain 
$ \startchain{\mathsf{r}_1}\chainedlink{\mathsf{a}_i}{\noact}\chainedlink{\noact}{\mathsf{a}_o}\chainedlink{\mathsf{b}_i}{\noact}\chainedlink{\noact}{\mathsf{b}_o}\chainend{\mathsf{r}_2}$.
